---
title: "Attitudinal Equity"
author: "Henrik Karlsson"
date: "16 mars 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set working directory where you're data is store.
```{r setwed}
setwd("G:/BU/Market_Science_Group/1.1 R/3. Development/AE")
```

### Required packages
```{r load packages, echo = TRUE, results="hide", message= FALSE}
library(haven)
library(tidyr)
library(dplyr)

```


### Preprocessing
Before the actual AE calculation, we need to prepare the data.

1. Read data and change all variable names to lower character
2. Rename the unique id to "uid" and control that number of unique ids = number of rows in the data file
3. Specify the beginning of the variable names for AE questions (relevance and closeness).
4. We need to make sure that all AE variable names begins equally for releveance and closeness and that the brand id is within the variable name, separated by a "_".

For example:

*Valid* variable names <- relevance for first brand = "q8aloop1_1", second brand = "q8aloop1_2" ... 

*Not* valid variable names <- closeness for first brand = "q8b_loop1_1".

There may only be one (1) "_" within the variable name for this script to work. However it can easly be changed if needed.


```{R preprocessing}
# Read data
r <- read_spss(file ="SE160810450101_170221_AE.sav")
names(r) <- tolower(names(r))

# AE question in data file
d <- rename(r, uid = respondent_serial)  # Replace unique respondent ID to uid

relevance <- "q8a"
closeness <- "q8b"

# Check unique id
length(unique(d$uid)) == nrow(d)

```

### Extract and arrange all relevance and closeness questions

```{r extract ae, warning = FALSE}
# Extract BRAND PERFORMANCE - Relevance

rel <- d %>%
        select(uid, starts_with(relevance)) %>%              # Select variables
        gather(variable, value, -uid, na.rm =T) %>%          # Transpose data
        separate(variable, into = c("question", "brand_id"), 
                 sep = "_", remove = T) %>%                  # split variable name to extract brand id
        select(-question, relevance = value) %>%             # remove question and rename value column
        mutate(brand_id = as.factor(brand_id)) %>%           # change data type of brand id
        arrange(uid, desc(brand_id))                         # sort data


# Extract BRAND CLOSENESS
clo <- d %>%
        select(uid, starts_with(closeness)) %>%
        gather(variable, value, -uid, na.rm =T) %>%
        separate(variable, into = c("question", "brand_id"), 
                 sep = "_", remove = T) %>%
        select(-question, closeness = value) %>%
        mutate(brand_id = as.factor(brand_id)) %>%
        arrange(uid, desc(brand_id))

```



### Compute AE

```{r ae}
# Combine PERFORMANCE and CLOSENESS and compute AE
ae <- left_join(x = rel, y = clo, by = c("uid", "brand_id")) %>%  # Merge rel and clo
        select(uid, brand_id, relevance, closeness) %>%           # Select necessary variables  
        mutate(avg_brand_score = (relevance + closeness)/2) %>%   # Compute average brand score
        group_by(uid) %>%                                         # Group data per respondent
        mutate(rank = rank(-avg_brand_score),                     # Rank avg brand score for each respondent
               step1 = 1/rank^1.35,                               # Compute numerator
               step2 = sum(step1),                                # Compute denominator  
               ae = step1/step2 * 100) %>%                        # Compute AE
        arrange(uid, desc(brand_id))

# Remove help calculations
ae <- ae %>%
        select(-avg_brand_score, -rank, -step1, -step2)

head(ae)

rm(rel, clo)

````
        



